---
title: "Lab1B"
author: "Michiel Klaassen"
date: "14-2-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vctrs)
library(zeallot)
```

```{r}
set.seed(123)
vec_1 <- rnorm(10)
vec_2 <- as.integer(vec_1)
vec_3 <- c(rep(TRUE, 3), c(rep(FALSE, 4)), NA)
vec_4 <- as.numeric(vec_3)
vec_5 <- c("Marc", "Ronald", "Maarten")
```

# 1. Creating some vectors
1. What does set.seed(123) do? Check the documentation!
    + Set a seed which makes sure the random values are constant from that point on
2. What does the above as.integer() function do?
    + Cast all numbers to whole numbers, integers.
3. And as.numeric()?
    + Transform the logical values "TRUE" and "FALSE" to 1 and 0 respectively.
4. Check the contents of vec_2, vec_3 and vec_4
```{r}
vec_2
vec_3
vec_4
```
5. Determine the class() of vec_1 to vec_5
```{r}
class(vec_1)
class(vec_2)
class(vec_3)
class(vec_4)
class(vec_5)
```
6. What happens to the data type if we combine (using c() vec_1 en vec_2)? Why?
    + It casts the integer vector to a number vector. It uses coercion and decides automatically up the tree of types what it can best cast to.
    + If we run ```vec_ptype2(vec_1, vec_2)``` we see that it suggests ```numeric(0)``` because that is what it will cast to. 
```{r}
vec_ptype2(vec_1, vec_2)
```
    + If we run ```vec_cast(vec_1, vec_2)``` it gives us an error that it can't convert doubles to integers due to loss of precision. Which is probably a hint why it gives us ```numeric(0)``` as suggestion: we won't lose any precision in either of the 2 vectors.
```{r, error=TRUE}
vec_cast(vec_1, vec_2)
```
7. What is the resulting datatype is we combine vec_4 and vec_5? Why?
    + Logically I would suggest that it would become a vector of characters. Because a number kind of keeps its meaning when tranforming it to a string, but converting a string to a number makes nu sense.
    + I am wondering why ```vec_ptype2``` gives me an error instead of suggesting the possible coercions.
```{r, error=TRUE}
c(vec_4, vec_5)
class(c(vec_4, vec_5))
vec_ptype2(vec_4, vec_5)
```
# 2. R as a calculator

1. What is the sum of 3, 12 and 250?
```{r}
sum(3, 12, 250)
```
2. What is the difference between 4902 and 3987?
```{r}
diff(c(4902, 3987))
```
3. Calculate 2 to the power of 24
```{r}
2^24
```
4. What is 240 divided by 14?
```{r}
240/14
```
5. What is the average of 12 and 20?
```{r}
mean(c(12, 20))
```
6. What is the average of 10.8, 13.7 and 15.9?
```{r}
mean(c(10.8, 13.7, 15.9))
```

# 3. Data types

1. Create a vector (`vec_8`) consisting of the numbers `10.5`, `20.4`, `30.5`, `28.9` and `5.4`.
```{r}
vec_8 <- c(10.5, 20.4, 30.5, 28.9, 5.4)
vec_8
```
2. Create a vector (`vec_9`) consisting of a decreasing sequence of integers from `50` to `-10`.
```{r}
vec_9 <- c(50:-10)
vec_9
```
3. Create a vector (`vec_10`) consisting of an increasing sequence of even integers from -`10` to `50.` You can use the `seq()` function here.
```{r}
vec_10 <- seq(from = -10, to = 50, by = 2)
vec_10
```
4. Add the number 5 to the tail of `vec_10`, and save the result as a new vector (`vec_11`).
```{r}
vec_11 <- c(vec_10, 5)
vec_11
```
5. Create a vector (`vec_12`) of five hundred alternating zeros and ones. Use the function rep() here.
```{r}
vec_12 <- rep(c(0, 1), 250)
vec_12
```
6. Create a vector (`vec_13`) of a hundred, uniformly distributed, random values between `-1` and `+1`. (Hint: see the function `runif()` ) Use `set.seed(123)` to set the random number generator.
```{r}
set.seed(123)
vec_13 <- runif(n = 100, min = -1, max = 1)
vec_13
```
7. Create a vector (`vec_14`) of hundred random normal distributed numbers with `mean = 2` and standard deviation is `0.3`. See `?rnorm()` for help.
```{r}
vec_14 <- rnorm(n = 100, mean = 2, sd = 0.3)
vec_14
```
8. Create a histogram (use function `hist()`) of vectors `vec_13` and `vec_14.` Can you notice something striking?
```{r}
hist(vec_13)
```
```{r}
hist(vec_14)
```
  
    + I'm not sure what the striking notice should be, but you can see that one is uniformly distributed and one is normally distributed?
9. Calculate the average of all values in `vec_14.` (Hint: see the function `mean()`. what did you expect to be the result? Why is the answer not exactly equal to `2`
```{r}
mean(vec_14)
```
    + You would expect the mean to be 2, but it's still random. There should be a seed available where (however unlikely it is) the mean is 1
10. Now replace position `5` in vector `vec_14` with `NA`, use `vec_14[5]` in you call. Confirm your operation on this vector.
```{r}
vec_15 <- vec_14
vec_15[5] = NA
vec_15
```
11. Calculate the mean of `vec_14` again. (Hint: make sure you deal with the missing values!). Look at `?mean()` for the solution.
```{r}
mean(vec_15, na.rm = TRUE)
```

# 4. Vector indices
1. Pull out all values that round down to 1.0 in vector `vec_1` *HINT use `as.integer()` than create a logical index, then use the logical index to get to your answer*
```{r}
round_to_one <- round(vec_1, digits = 0) == 1
round_to_one
vec_1[round_to_one]
round(vec_1, digits = 0)
```
    + Actually rounding doesn't yield any results as your can see. But when we use `as.integer`, then sure.
```{r}
as_int <- as.integer(vec_1) == 1
as_int
vec_1[as_int]
```

# 5. The Matrix

1. Create a matrix (matrix_1) of four columns and three rows with increasing integers from 1 to 12, left to right and then top to bottom. Use the function matrix(). See the help first.
```{r}
matrix_data <- c(1:12)
matrix_1 <- matrix(matrix_data, 3, 4, byrow = TRUE)
matrix_1
```
2. Create a matrix (matrix_2) of four columns and three rows, the first column only containing the value 1, the second only 2, the third only 3 and the fourth column only 4.
```{r}
matrix_2_data <- c(1:4)
matrix_2 <- matrix(matrix_2_data, 3, 4, byrow = TRUE)
matrix_2
```
3. Calculate the sum of matrix_1. (Hint: see the function sum())
```{r}
sum(matrix_1)
```

# 6. From vectors to tibbles

1. A numerical vector named `numbers` consisting of the following numbers: `1.8, 4.5, 10.1, 8.3`.
```{r}
numbers <- c(1.8, 4.5, 10.1, 8.3)
numbers
```
2. A numerical vector named `prime` consisting of the following integers: `1, 3, 5, 7` and `11`
```{r}
prime <- c(1, 3, 5, 7, 11)
prime
```
3. A character vector named `crazy` consisting of the following values: `‘abc1’, ‘foo2’, ‘bar3’, ‘app5’ `and` ‘yepp7’`
```{r}
crazy <- c("abc1", "foo2", "bar3", "app5", "yepp7")
crazy
```
4. A logical vector named `valid` consisting of: `TRUE, FALSE, FALSE, TRUE` and `FALSE`
```{r}
valid <- c(TRUE, FALSE, FALSE, TRUE, FALSE)
valid
```
5. Try to create a list called `mylist` containing the vectors numbers, prime, crazy and valid.
```{r}
mylist <- c(numbers, prime, crazy, valid)
mylist
```
6. Try to create a tibble named `mydata` with the function `tibble()` from the above four vectors. This will fail, why? .
```{r, error=TRUE}
mydata <- tibble(numbers, prime, crazy, valid)
mydata
```
    + It fails because it doesn't have the same amount of values in each vector.
7. Fix the error you found by creating an extra `NA` value in one of the vectors. Create the dataframe. Also try giving the columns different names. (Hint: see the function `names()` and the function `tibble()`.
```{r}
mydata_fixed <- tibble(c(numbers, NA), prime, crazy, valid)
names(mydata_fixed) <- c("numbers", "prime", "crazy", "valid")
mydata_fixed
```
8. What happens if you try to calculate the sum of `mydata`?
```{r, error=TRUE}
sum(mydata_fixed)
```
    + Error, they're not all numbers.
9. When you look at the structure of the tibble using `str()`, you will see that the crazy column with the `“abc1”, “foo2”`, etc values has a certain data type. Which data type is “`name`”? Do you think it is the right type? Or do you need to change it? To what other type could you change it?
```{r}
str(mydata_fixed)
```
    + I'm not so sure what the question is for this?

