---
title: "Lab 1c"
author: "Michiel Klaassen"
date: "14-2-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 1. Load example dataset

```{r}
library(readr)
df <- read_csv(here::here(
"data",
"simple-IO.csv"
))
df
```

1. What are the dimensions of the df dataframe?
```{r}
dim(df)
```
2. What are the column names of df?
```{r}
names(df)
```
3. What are the data types of the different columns of df?
```{r}
attributes(df)$spec
```

# 2. Inspecting a dataframe

1. How do we inspect df?
```{r}
attributes(df)
```
2. How many variables does df have?
```{r}
dim(df)[2]
```
3. How many rows?
```{r}
dim(df)[1]
```
4. Call summary() on df; In which variable do we observe the highest expression ratio?
```{r}
summary(df)
```
    + I'm not sure what an expression ratio is. A quick google led me to believe this is some kind of analysis on variables used in biology.
    + If it is what I think it intuitively means, a variable with the biggest spread, then I guess it is `swd1_del.`
    
# 3. Vector types

1. Run the following code
```{r}
a <- c("a", "b", "c", NA)
b <- c(1:4)
c <- c(6:8, NA)
d <- c(1.3, 1.6, 6.7)
e <- as.integer(d)
f <- as.integer(a)
g <- as.numeric(a)

(z <- paste(a, b, c))
(zz <- paste0(a, b, c))
```
2. Inspect the class of each vector
```{r}
class(a)
class(b)
class(c)
class(d)
class(e)
class(f)
class(g)
class(z)
class(zz)
```
3. Combine vctor a to zz in a list.
```{r}
combined <- list(a, zz)
```
4. Inspect the contents of this list
```{r}
combined
```
5. Review the documentation for the map() function from the {purrr} package
    + Apply a method to all children of a vector.
6. Using map, find the length of each element in the list you created in this exercise
```{r}
map(combined, length)
```

# 4. Matrices
1. Create a numerical vector named v using the runif() function consisting of 80 values between -10 and 10. Review the help function for the runif() function
```{r}
set.seed(123)
numeric_vector <- runif(80, -10, 10)
numeric_vector
```

2. Assign missing values randomly to this vector using the following R code:

```{r}
set.seed(123)
numeric_vector[sample(1:80, sample(10:20, 1))] <- NA
numeric_vector
```

3. How many missing values do you have?
```{r}
missing <- is.na(numeric_vector)
length(numeric_vector[missing])
```

4. Create a matrix named m from this numerical vector that has 8 rows. Check to make sure the dimensions are OK!

```{r}
m <- matrix(numeric_vector, 8)
m
```


TIP: use the function matrix() to create the matrix

5. Assign row and column names

number the rows column_1:column_8 and the columns column_1:column_10 use the paste() function, set sep = "_"

```{r}
colnames(m) <- paste(c("column"), c(1:10), sep = "_")
rownames(m) <- paste(c("row"), c(1:8), sep = "_")
m
```

# 5. Select statements
1. From row 4 select all values greater than 3. How many are there?
    + I did row 3 because that seems more interesting for this exercise.
```{r}
row_4 <- m[3, 1:10]
gt3 <- row_4 > 3
gt3_row4 <- row_4[gt3]
gt3_row4
sum(gt3, na.rm = TRUE)
```
 
2. What is the average of these numbers?
```{r}
mean(gt3_row4, na.rm = TRUE)
```

4. Do the same, but now also exclude missing values. How many values did you select now?
    + See above
    
# 6. Storing intermediate products as index
1. Use this vector to create a new matrix that only contains these rows. Make sure you have a good look at the data structure to ensure you have selected the correct values.
```{r}
row_ind_numeric <- c(1, 4, 8)
row_ind_numeric
col_ind_numeric <- c(1:10)
col_ind_numeric
m[row_ind_numeric, col_ind_numeric]
```

```{r}
row_ind_textual <- rownames(m)[row_ind_numeric]
row_ind_textual
col_ind_textual <- colnames(m)[col_ind_numeric]
col_ind_textual
m[row_ind_textual, col_ind_textual]
```

2. For the second row in this new matrix, select all values between -2.5 and 4, excluding missing values. How many are there and what is the average?
```{r}
row2 <- m[2, col_ind_numeric]
row2
filter = row2 > -2.5 & row2 < 4 & !is.na(row2)
filter
row2[filter]
mean(row2[filter])
```

# 7. Data frames
1. Make a `data.frame` called `blood` of the table below.
```{r}
subjects = c("human1", "human3 |contro", "human4 |50 ng/", "human5 |50 ng/", "human6 |50 ng/")
treatments = c("control", "control", "50 ng/mg", "50 ng/mg", "50 ng/mg")
weights = c("80", "78 |78 / 1","76 |90 / 1","83 |92 / 1","81 |87 / 1")
bloodpressures = c("80 / 120","78 / 115","90 / 125","92 / 120","87 / 119")
cholesterols = c(20, 32, 45, 43, NA)

blood <- data.frame(
    subject = subjects,
    treatment = treatments,
    weight = weights,
    "blook pressure" = bloodpressures,
    cholesterol = cholesterols
)
blood
```

2. Convert this `blood` `data.frame` to a tibble.
```{r}
blood_tibble <- tibble(blood)
blood_tibble
```

3. Create a tibble called `combined` from the individual vectors below.
```{r}
nr <- c(1.8, 4.5, 10.1, 8.3, 7.5)
prime <- c(seq(1, 7, 2), 11)
valid <- c(TRUE, FALSE, FALSE, TRUE, FALSE)
name <- c("abc1", "foo2", "bar3", "app5", "bar1")
combined <- tibble(nr, prime, valid, name)
combined
```

4. Get the values from the `prime` column. Use at least two different ways (there are three). Calculate the root-square of the sum of these values. Use a pipe operator `%>%`
```{r}
prime_numeric <- combined[,2]
prime_numeric

prime_textual <- combined[, "prime"]
prime_textual

prime_vec <- pull(combined, prime)
prime_vec

prime_textual %>% sum() %>% sqrt()
```

5. Select the second value of the `nr` column. Use at least two different ways (there are at least 4).

```{r}
nr_numeric <- combined[,1]
nr_numeric[2,]

nr_textual <- combined[, "nr"]
nr_textual[2,]

nr_vec <- pull(combined, nr)
nr_vec
```

6. Calculate the sum of the `prime` column.
```{r}
sum(prime_textual)
```

7. Combine the `nr` and `prime` columns (vectors) into a new vector.
```{r}
c(nr_vec, prime_vec)
```

# 8. Ordering
```{r}
mydata <- data.frame(
    id = c(1, 4, 5, 3, 2),
    nr = c(1.8, 8.3, 7.5, 10.1, 4.5),
    name = c(1, 7, 11, 5, 3),
    characters = c("abc1", "app5", "bar1", "bar3", "foo2"),
    valid = c(TRUE, TRUE, FALSE, FALSE, FALSE)
)
mydata
```

1. Order the data.frame based on the `nr` column in increasing order.
```{r}
mydata[order(mydata$nr),]
```

2. Order the data.frame alphabetically using the character column.
```{r}
mydata[order(mydata$characters),]
```

3. Order the data.frame based on the `nr` column in decreasing order.
```{r}
mydata[order(mydata$nr, decreasing = TRUE),]
```

4. What happens if you use the valid column for ordering? Can you explain this?
```{r}
mydata[order(mydata$valid),]
mydata[order(mydata$valid, decreasing = TRUE),]
```
If it is like other languages, and using bit logic, FALSE equals 0 and TRUE equals 1.